A pipeline to analyse time-course gene expression data

### 2019-11-19 manuscript.tex
Deploy NelleV/2019timecourse-rnaseq-pipeline to github.com/NelleV/2019timecourse-rnaseq-pipeline.git:gh-pages

The manuscript was last compile on 2019-11-13 08:34:55.

@@

@@

\begin{center}\includegraphics{reports/manuscript_files/figure-latex/unnamed-chunk-9-1} \end{center}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{kable}\NormalTok{(}\KeywordTok{head}\NormalTok{(moanin_model}\OperatorTok{$}\NormalTok{basis))}
\end{Highlighting}
\end{Shaded}

\begin{tabular}{l|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r}
\hline
  & GroupM & GroupK & GroupC & GroupVL & GroupVH & GroupM:splines::ns(Timepoint, df = degrees\_of\_freedom)1 & GroupK:splines::ns(Timepoint, df = degrees\_of\_freedom)1 & GroupC:splines::ns(Timepoint, df = degrees\_of\_freedom)1 & GroupVL:splines::ns(Timepoint, df = degrees\_of\_freedom)1 & GroupVH:splines::ns(Timepoint, df = degrees\_of\_freedom)1 & GroupM:splines::ns(Timepoint, df = degrees\_of\_freedom)2 & GroupK:splines::ns(Timepoint, df = degrees\_of\_freedom)2 & GroupC:splines::ns(Timepoint, df = degrees\_of\_freedom)2 & GroupVL:splines::ns(Timepoint, df = degrees\_of\_freedom)2 & GroupVH:splines::ns(Timepoint, df = degrees\_of\_freedom)2 & GroupM:splines::ns(Timepoint, df = degrees\_of\_freedom)3 & GroupK:splines::ns(Timepoint, df = degrees\_of\_freedom)3 & GroupC:splines::ns(Timepoint, df = degrees\_of\_freedom)3 & GroupVL:splines::ns(Timepoint, df = degrees\_of\_freedom)3 & GroupVH:splines::ns(Timepoint, df = degrees\_of\_freedom)3 & GroupM:splines::ns(Timepoint, df = degrees\_of\_freedom)4 & GroupK:splines::ns(Timepoint, df = degrees\_of\_freedom)4 & GroupC:splines::ns(Timepoint, df = degrees\_of\_freedom)4 & GroupVL:splines::ns(Timepoint, df = degrees\_of\_freedom)4 & GroupVH:splines::ns(Timepoint, df = degrees\_of\_freedom)4 & GroupM:splines::ns(Timepoint, df = degrees\_of\_freedom)5 & GroupK:splines::ns(Timepoint, df = degrees\_of\_freedom)5 & GroupC:splines::ns(Timepoint, df = degrees\_of\_freedom)5 & GroupVL:splines::ns(Timepoint, df = degrees\_of\_freedom)5 & GroupVH:splines::ns(Timepoint, df = degrees\_of\_freedom)5 & GroupM:splines::ns(Timepoint, df = degrees\_of\_freedom)6 & GroupK:splines::ns(Timepoint, df = degrees\_of\_freedom)6 & GroupC:splines::ns(Timepoint, df = degrees\_of\_freedom)6 & GroupVL:splines::ns(Timepoint, df = degrees\_of\_freedom)6 & GroupVH:splines::ns(Timepoint, df = degrees\_of\_freedom)6\\
\hline
GSM1557140 & 0 & 1 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0\\
\hline
GSM1557141 & 0 & 1 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0\\
\hline
GSM1557142 & 0 & 1 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0 & 0 & 0.0000000 & 0 & 0 & 0\\
\hline
GSM1557143 & 0 & 1 & 0 & 0 & 0 & 0 & 0.5142857 & 0 & 0 & 0 & 0 & 0.0357143 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & -0.1546016 & 0 & 0 & 0 & 0 & 0.3478535 & 0 & 0 & 0 & 0 & -0.1932519 & 0 & 0 & 0\\
\hline
GSM1557144 & 0 & 1 & 0 & 0 & 0 & 0 & 0.5142857 & 0 & 0 & 0 & 0 & 0.0357143 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & -0.1546016 & 0 & 0 & 0 & 0 & 0.3478535 & 0 & 0 & 0 & 0 & -0.1932519 & 0 & 0 & 0\\
\hline
GSM1557145 & 0 & 1 & 0 & 0 & 0 & 0 & 0.5142857 & 0 & 0 & 0 & 0 & 0.0357143 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & -0.1546016 & 0 & 0 & 0 & 0 & 0.3478535 & 0 & 0 & 0 & 0 & -0.1932519 & 0 & 0 & 0\\
\hline
\end{tabular}

@@

Another difficulty with such an approach is making sense of the general
temporal structure for any particular gene, as different genes will have
different combinations of timepoints DE. For the comparison of the Kawasaki
strain to the control, for example, there are \texttt{length(signifCombos)} genes
found DE in some timepoints, and there are \texttt{length(tabCombos)} different
combinations of timepoints for which they are DE. Some of these make sense,
such as DE in timepoints 48H-168H (\texttt{tabCombos{[}"48,60,72,120,168"{]}} genes), but
many are very fragmentary. For example there are \texttt{tabCombos{[}"48,60,120,168"{]}}
genes which are DE in timepoints 48,60,120,168H, but \emph{not} in the 72H. Many of
these genes are likely to have not made the cutoff for significance in 72H,
but don't show real differences in the overall trend between 48H-168H.

@@

\begin{longtable}[]{@{}lll@{}}
\toprule
Name & Formula &\tabularnewline
\midrule
\endhead
timely & \(\text{lfc}(t)\) & Function of time\tabularnewline
sum & \(\sum_t lfc(t)\) & Sum of log fold change.\tabularnewline
abs\_sum & \(\sum_t \lvert lfc(t)\lvert\) & Always positive\tabularnewline
max & \(\max_t \lvert lfc(t) \lvert\) & Always positive\tabularnewline
min & \(\min_t \lvert lfc(t) \lvert\) & Always positive\tabularnewline
timecourse & See details below & Captures overall strength and direction\tabularnewline
\bottomrule
\end{longtable}

@@

\[\text{sign}\left(\frac{1}{T}\sum_{t = 1}^T lfc(t) \right) \times \left(\frac{1}{T}\sum_{t= 1}^T \lvert {lfc(t)} \lvert \right)\]

@@

@@

\KeywordTok{plot}\NormalTok{(lfc_timecourse, }\OperatorTok{-}\KeywordTok{log10}\NormalTok{(pvalue), }\DataTypeTok{pch=}\DecValTok{20}\NormalTok{, }\DataTypeTok{main=}\StringTok{"Volcano plot"}\NormalTok{,}
     \DataTypeTok{xlim=}\KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{2.5}\NormalTok{, }\DecValTok{2}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{#<c2><a0>Then assign scores and labels to all the data, using a goodness-of-fit}
\CommentTok{# scoring function.}
\NormalTok{scores_and_labels =}\StringTok{ }\KeywordTok{splines_kmeans_score_and_label}\NormalTok{(}
\NormalTok{    data, kmeans_clusters)}
\NormalTok{labels =}\StringTok{ }\NormalTok{scores_and_labels}\OperatorTok{$}\NormalTok{labels}

\CommentTok{# Let's keep only the list of genes that have a label.}
\NormalTok{scores =}\StringTok{ }\KeywordTok{rowMin}\NormalTok{(scores_and_labels}\OperatorTok{$}\NormalTok{scores[}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(labels), ])}
\NormalTok{labels =}\StringTok{ }\KeywordTok{unlist}\NormalTok{(labels[}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(labels)])}
\NormalTok{genes_in_cluster2 =}\StringTok{ }\KeywordTok{names}\NormalTok{(labels[labels}\OperatorTok{==}\NormalTok{cluster_of_interest])}
\end{Highlighting}
\end{Shaded}

Let us now, for each cluster, visualize the distribution of scores of the
genes assigned to that cluster.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Get the best score for each gene}
\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow=}\KeywordTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{, }\DecValTok{3}\NormalTok{))}
\NormalTok{n_clusters =}\StringTok{ }\KeywordTok{dim}\NormalTok{(kmeans_clusters}\OperatorTok{$}\NormalTok{centroids)[}\DecValTok{1}\NormalTok{]}
\ControlFlowTok{for}\NormalTok{(cluster_id }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{n_clusters)\{}
    \KeywordTok{hist}\NormalTok{(scores[labels }\OperatorTok{==}\StringTok{ }\NormalTok{cluster_id],}
     \DataTypeTok{breaks=}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{50}\OperatorTok{/}\DecValTok{50}\NormalTok{), }\DataTypeTok{xlim=}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{),}
     \DataTypeTok{col=}\StringTok{"black"}\NormalTok{, }\DataTypeTok{main=}\KeywordTok{paste}\NormalTok{(}\StringTok{"C"}\NormalTok{, cluster_id, }\DataTypeTok{sep=}\StringTok{""}\NormalTok{),}
     \DataTypeTok{xlab=}\StringTok{"score"}\NormalTok{, }\DataTypeTok{ylab=}\StringTok{"Num. genes"}\NormalTok{)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

@@

\begin{center}\includegraphics{reports/manuscript_files/figure-latex/unnamed-chunk-28-1} \end{center}

\textbf{EAP: I don't know what to think about these plots -- these are best scoring,
but pretty noisy. Also the per-gene spline is sometimes not as good of a
explanation as the cluster centroid. It's seems like an important plot to
make}-

\textbf{NV: this were the worst scoring genes. I fixed this and the plots now look
much better}

@@

\begin{center}\includegraphics{reports/manuscript_files/figure-latex/unnamed-chunk-31-1} \end{center}

@@

\begin{center}\includegraphics{reports/manuscript_files/figure-latex/unnamed-chunk-36-1} \end{center}

@@

\begin{center}\includegraphics{reports/manuscript_files/figure-latex/unnamed-chunk-36-2} \end{center}

@@

@@

\begin{center}\includegraphics{reports/manuscript_files/figure-latex/unnamed-chunk-39-1} \end{center}

@@

\textbf{EAP: \texttt{plot\_model\_explorer} is relatively slow, because it does all those pairwise calculations. And then to do it for smaller subset, redo all the calculations. Can't the output be saved, or the calculations be a different step from the plotting?}

>>>>

The manuscript was last compile on 2019-11-19 02:22:00.

@@

\hypertarget{package-versions}{%
\subsection{Package versions}\label{package-versions}}

The following packages are needed for this workflow.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# From CRAN}
\KeywordTok{library}\NormalTok{(NMF)}
\KeywordTok{library}\NormalTok{(ggfortify)}

\CommentTok{# From Bioconductor}
\KeywordTok{library}\NormalTok{(topGO)}
\KeywordTok{library}\NormalTok{(biomaRt)}

\CommentTok{# From GitHub}
\KeywordTok{library}\NormalTok{(moanin)}
\end{Highlighting}
\end{Shaded}

To install \texttt{moanin} directly from GitHub, use the devtools' \texttt{install\_github}
function:

@@

\begin{center}\includegraphics{reports/manuscript_files/figure-latex/unnamed-chunk-8-1} \end{center}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{kable}\NormalTok{(}\KeywordTok{head}\NormalTok{(moanin_model}\OperatorTok{$}\NormalTok{basis), }\DataTypeTok{digits=}\DecValTok{2}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{font_size =} \DecValTok{7}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[H]
\centering\begingroup\fontsize{7}{9}\selectfont

\begin{tabular}{l|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r}
\hline
  & GroupM & GroupK & GroupC & GroupVL & GroupVH & GroupM:splines::ns(Timepoint, df = degrees\_of\_freedom)1 & GroupK:splines::ns(Timepoint, df = degrees\_of\_freedom)1 & GroupC:splines::ns(Timepoint, df = degrees\_of\_freedom)1 & GroupVL:splines::ns(Timepoint, df = degrees\_of\_freedom)1 & GroupVH:splines::ns(Timepoint, df = degrees\_of\_freedom)1 & GroupM:splines::ns(Timepoint, df = degrees\_of\_freedom)2 & GroupK:splines::ns(Timepoint, df = degrees\_of\_freedom)2 & GroupC:splines::ns(Timepoint, df = degrees\_of\_freedom)2 & GroupVL:splines::ns(Timepoint, df = degrees\_of\_freedom)2 & GroupVH:splines::ns(Timepoint, df = degrees\_of\_freedom)2 & GroupM:splines::ns(Timepoint, df = degrees\_of\_freedom)3 & GroupK:splines::ns(Timepoint, df = degrees\_of\_freedom)3 & GroupC:splines::ns(Timepoint, df = degrees\_of\_freedom)3 & GroupVL:splines::ns(Timepoint, df = degrees\_of\_freedom)3 & GroupVH:splines::ns(Timepoint, df = degrees\_of\_freedom)3 & GroupM:splines::ns(Timepoint, df = degrees\_of\_freedom)4 & GroupK:splines::ns(Timepoint, df = degrees\_of\_freedom)4 & GroupC:splines::ns(Timepoint, df = degrees\_of\_freedom)4 & GroupVL:splines::ns(Timepoint, df = degrees\_of\_freedom)4 & GroupVH:splines::ns(Timepoint, df = degrees\_of\_freedom)4 & GroupM:splines::ns(Timepoint, df = degrees\_of\_freedom)5 & GroupK:splines::ns(Timepoint, df = degrees\_of\_freedom)5 & GroupC:splines::ns(Timepoint, df = degrees\_of\_freedom)5 & GroupVL:splines::ns(Timepoint, df = degrees\_of\_freedom)5 & GroupVH:splines::ns(Timepoint, df = degrees\_of\_freedom)5 & GroupM:splines::ns(Timepoint, df = degrees\_of\_freedom)6 & GroupK:splines::ns(Timepoint, df = degrees\_of\_freedom)6 & GroupC:splines::ns(Timepoint, df = degrees\_of\_freedom)6 & GroupVL:splines::ns(Timepoint, df = degrees\_of\_freedom)6 & GroupVH:splines::ns(Timepoint, df = degrees\_of\_freedom)6\\
\hline
GSM1557140 & 0 & 1 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0\\
\hline
GSM1557141 & 0 & 1 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0\\
\hline
GSM1557142 & 0 & 1 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0 & 0 & 0.00 & 0 & 0 & 0\\
\hline
GSM1557143 & 0 & 1 & 0 & 0 & 0 & 0 & 0.51 & 0 & 0 & 0 & 0 & 0.04 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & -0.15 & 0 & 0 & 0 & 0 & 0.35 & 0 & 0 & 0 & 0 & -0.19 & 0 & 0 & 0\\
\hline
GSM1557144 & 0 & 1 & 0 & 0 & 0 & 0 & 0.51 & 0 & 0 & 0 & 0 & 0.04 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & -0.15 & 0 & 0 & 0 & 0 & 0.35 & 0 & 0 & 0 & 0 & -0.19 & 0 & 0 & 0\\
\hline
GSM1557145 & 0 & 1 & 0 & 0 & 0 & 0 & 0.51 & 0 & 0 & 0 & 0 & 0.04 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & -0.15 & 0 & 0 & 0 & 0 & 0.35 & 0 & 0 & 0 & 0 & -0.19 & 0 & 0 & 0\\
\hline
\end{tabular}
\endgroup{}
\end{table}

@@

Another difficulty with such an approach is making sense of the general
temporal structure for any particular gene, as different genes will have
different combinations of timepoints DE. For the comparison of the Kawasaki
strain to the control, for example, there are 26534 genes
found DE in some timepoints, and there are 1590 different
combinations of timepoints for which they are DE. Some of these make sense,
such as DE in timepoints 48H-168H (509 genes), but
many are very fragmentary. For example there are 330
genes which are DE in timepoints 48,60,120,168H, but \emph{not} in the 72H. Many of
these genes are likely to have not made the cutoff for significance in 72H,
but don't show real differences in the overall trend between 48H-168H.

@@

\begin{longtable}[]{@{}lll@{}}
\toprule
Name & Formula &\tabularnewline
\midrule
\endhead
timely & \(\text{lfc}(t)\) & Function of time\tabularnewline
sum & \(\sum_t \text{lfc}(t)\) & Sum of log fold change.\tabularnewline
abs\_sum & \(\sum_t \lvert \text{lfc}(t)\lvert\) & Always positive\tabularnewline
max & \(\max_t \lvert \text{lfc}(t) \lvert\) & Always positive\tabularnewline
min & \(\min_t \lvert \text{lfc}(t) \lvert\) & Always positive\tabularnewline
timecourse & See details below & Captures overall strength and direction\tabularnewline
\bottomrule
\end{longtable}

@@

\[\text{sign}\left(\frac{1}{T}\sum_{t = 1}^T \text{lfc}(t) \right) \times \left(\frac{1}{T}\sum_{t= 1}^T \lvert {\text{lfc}(t)} \lvert \right)\]

@@

The returning object is a matrix, where each row corresponds to a gene,
each column to a contrasts, and each entry to the log-fold change for this
pair of contrast and gene.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{kable}\NormalTok{(log_fold_change_timecourse[}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{, ], }\DataTypeTok{digits=}\DecValTok{3}\NormalTok{, }\DataTypeTok{booktabs=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{tabular}{lrrrr}
\toprule
  & M-K & M-C & M-VL & M-VH\\
\midrule
NM\_009912 & 0.432 & 0.753 & -0.836 & -0.488\\
NM\_008725 & -1.257 & -2.008 & 1.924 & 1.729\\
NM\_007473 & 0.385 & 0.335 & 0.516 & 0.263\\
ENSMUST00000094955 & 0.355 & 0.376 & -0.316 & 0.256\\
NM\_001042489 & -0.266 & 0.410 & 0.415 & 0.374\\
\bottomrule
\end{tabular}

Each methods of computing the log-fold change captures different elements of
the time-course data: overall change, largest change, \ldots{}

\begin{center}\includegraphics{reports/manuscript_files/figure-latex/unnamed-chunk-19-1} \end{center}

@@

\KeywordTok{plot}\NormalTok{(lfc_timecourse, }\OperatorTok{-}\KeywordTok{log10}\NormalTok{(pvalue), }\DataTypeTok{pch=}\DecValTok{20}\NormalTok{, }\DataTypeTok{main=}\StringTok{"Volcano plot"}\NormalTok{,}
     \DataTypeTok{xlim=}\KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{2.5}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{xlab=}\StringTok{"Timecourse lfc"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

The \texttt{splines\_kmeans} function returns a named list with:

\begin{itemize}
\tightlist
\item
  \texttt{centroids}: a matrix containing the cluster centroids. The matrix is of
  shape (\texttt{n\_centroid}, \texttt{n\_samples}).
\item
  \texttt{clusters}: a vector of size \texttt{n\_genes}, containing the
\end{itemize}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{#<c2><a0>Then assign scores and labels to all the data, using a goodness-of-fit}
\CommentTok{# scoring function.}
\NormalTok{scores_and_labels =}\StringTok{ }\KeywordTok{splines_kmeans_score_and_label}\NormalTok{(}
\NormalTok{    data, kmeans_clusters)}
\end{Highlighting}
\end{Shaded}

The \texttt{scores\_and\_labels} list contains two elements:

\begin{verbatim}
- `scores_and_labels$scores`: the matrix of shape `n_cluster` x `n_genes`,
  containing for each gene and each cluster, the goodness of fit as
  described above.
- `scores_and_labels$labels`: the labels for all of the genes with a
  good-enough goodness-of-fit score.
\end{verbatim}

First, let us visualize the distribution of goodness-of-fit scores for each
cluster. Here, we display the whole set of scores assigned to each cluster,
prior to filtering poor-quality fits.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Get the best score and best label for all of the genes}
\NormalTok{scores =}\StringTok{ }\KeywordTok{rowMin}\NormalTok{(scores_and_labels}\OperatorTok{$}\NormalTok{scores)}
\NormalTok{labels =}\StringTok{ }\KeywordTok{apply}\NormalTok{(scores_and_labels}\OperatorTok{$}\NormalTok{scores, }\DecValTok{1}\NormalTok{, which.min)}
\NormalTok{max_score =}\StringTok{ }\KeywordTok{quantile}\NormalTok{(scores, }\FloatTok{0.5}\NormalTok{)}
\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow=}\KeywordTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{3}\NormalTok{))}
\NormalTok{n_clusters =}\StringTok{ }\KeywordTok{dim}\NormalTok{(kmeans_clusters}\OperatorTok{$}\NormalTok{centroids)[}\DecValTok{1}\NormalTok{]}
\ControlFlowTok{for}\NormalTok{(cluster_id }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{n_clusters)\{}
    \KeywordTok{hist}\NormalTok{(scores[labels }\OperatorTok{==}\StringTok{ }\NormalTok{cluster_id],}
     \DataTypeTok{breaks=}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{50}\OperatorTok{/}\DecValTok{50}\NormalTok{), }\DataTypeTok{xlim=}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{),}
     \DataTypeTok{col=}\StringTok{"black"}\NormalTok{, }\DataTypeTok{main=}\KeywordTok{paste}\NormalTok{(}\StringTok{"C"}\NormalTok{, cluster_id, }\DataTypeTok{sep=}\StringTok{""}\NormalTok{),}
     \DataTypeTok{xlab=}\StringTok{"score"}\NormalTok{, }\DataTypeTok{ylab=}\StringTok{"Num. genes"}\NormalTok{)}
    \KeywordTok{abline}\NormalTok{(}\DataTypeTok{v=}\NormalTok{max_score, }\DataTypeTok{col=}\StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd=}\DecValTok{3}\NormalTok{, }\DataTypeTok{lty=}\DecValTok{2}\NormalTok{)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

Note that for some of the genes, the best scores is 1. A score of 1 indicates
that there is no correlation between the gene and the cluster: those genes fit
poorly to all clusters and the assignment to one cluster is done randomly..
This underlines the importance of filtering genes that fit poorly to clusters.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{labels =}\StringTok{ }\NormalTok{scores_and_labels}\OperatorTok{$}\NormalTok{labels}

\CommentTok{# Let's keep only the list of genes that have a label.}
\NormalTok{labels =}\StringTok{ }\KeywordTok{unlist}\NormalTok{(labels[}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(labels)])}

\CommentTok{# And also keep track of all the genes in cluster 2.}
\NormalTok{genes_in_cluster2 =}\StringTok{ }\KeywordTok{names}\NormalTok{(labels[labels}\OperatorTok{==}\NormalTok{cluster_of_interest])}
\end{Highlighting}
\end{Shaded}

@@

@@

\begin{center}\includegraphics{reports/manuscript_files/figure-latex/unnamed-chunk-30-1} \end{center}

@@

\begin{center}\includegraphics{reports/manuscript_files/figure-latex/unnamed-chunk-35-1} \end{center}

@@

\begin{center}\includegraphics{reports/manuscript_files/figure-latex/unnamed-chunk-35-2} \end{center}

@@

\begin{center}\includegraphics{reports/manuscript_files/figure-latex/unnamed-chunk-37-1} \end{center}

@@

@@
